### 2020-09-08 Latex/tehran-thesis - Copy/tex/enTitle.tex

% !TeX root=../main.tex
% در این فایل، عنوان پایان‌نامه، مشخصات خود و چکیده پایان‌نامه را به انگلیسی، وارد کنید.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\latinuniversity{University of Tehran}
\latincollege{College of Engineering}
\latinfaculty{Faculty of ٍElectrical and Computer Engineering}
\latindepartment{Network department}
\latinsubject{ٍElectrical Engineering}
\latinfield{Communication and Network}
\latintitle{Joint Power Allocation and Network Slicing in an
End-to-End ORAN System}
\firstlatinsupervisor{Dr. Shahmansouri}
%\secondlatinsupervisor{Second Supervisor}
%\firstlatinadvisor{First Advisor}
%\secondlatinadvisor{Second Advisor}
\latinname{Mojdeh}
\latinsurname{Karbalaee Motalleb}
\latinthesisdate{December 2019}
\latinkeywords{Writing Thesis, Template, \LaTeX, \XePersian}
\en-abstract{
This thesis studies on writing projects, theses and dissertations using tehran-thesis class. It ...
}