### 2020-09-10 Latex/tehran-thesis - Copy/tex/enTitle.tex
abstract add to thesis

% !TeX root=../main.tex
% در این فایل، عنوان پایان‌نامه، مشخصات خود و چکیده پایان‌نامه را به انگلیسی، وارد کنید.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\latinuniversity{University of Tehran}
\latincollege{College of Engineering}
\latinfaculty{Faculty of ٍElectrical and Computer Engineering}
\latindepartment{Network department}
\latinsubject{ٍElectrical Engineering}
\latinfield{Communication and Network}
\latintitle{Joint Power Allocation and Network Slicing in an
End-to-End ORAN System}
\firstlatinsupervisor{Dr. Shahmansouri}
%\secondlatinsupervisor{Second Supervisor}
%\firstlatinadvisor{First Advisor}
%\secondlatinadvisor{Second Advisor}
\latinname{Mojdeh}
\latinsurname{Karbalaee Motalleb}
\latinthesisdate{September 2020}
\latinkeywords{Writing Thesis, Template, \LaTeX, \XePersian}
\en-abstract{
Open radio access network (ORAN) alliance which has been formed recently establishes a flexible,  open, and smart radio access network (RAN) by combing the ideas from xRAN and cloud RAN (C-RAN). ORAN divides the functions of the RAN into three parts, namely remote unit (RU), distributed unit (DU), and central unit (CU). While RU contains lower PHY functions, DU contains higher PHY, MAC, and RLC and CU contains RRC, PDCP, and SDAP. CU and DU are implemented as virtual network functions (VNFs) running on a cloud environment. Interfaces between RU, CU, and DU are open standard interfaces. Network slicing as a new concept in 5G systems is used to share the network resources between various services while the operation of one service does not affect another service. In this paper, we study the problem of RAN network slicing in an ORAN system.  We formulate the problem of wireless link scheduling, assigning the slices to the services, and assigning the physical data centers resource to slices which is 3D-bin packing problems. The objective is to jointly maximize the energy efficiency and minimize power consumption of RUs and the cost of physical resources in a downlink channel. The problem is formulated as a mixed-integer optimization problem that can be decomposed into two independent sub-problems. Heuristic algorithms are proposed for each of the sub-problems.
}